Untwisting the Tetrahelix (v0.5)

### 2017-02-17 doc/TetrahelixUntwist.tex
significant corrections

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newtheorem{conjecture}{Conjecture}
\newtheorem{proofsketch}{Proof Sketech}
\usepackage{draftwatermark}

@@

\title{Untwisting the Tetrahelix (v0.4)}
\author{Robert L. Read \texttt{read.robert@gmail.com} \and
  Robert Gatliff \texttt{robert@toubat.org}
}

@@

\begin{align*}
  \text{one-hop} &= dist(H_{general}(n,c,\rho,d_{\rho}),H_{general}(n,c+1,\rho,d_{\rho}),r_{\rho})  \\
  \text{one-hop} &= dist(H_{general}(0,0,\rho,d_{\rho}),H_{general}(0,1,\rho,d_{\rho},r_{\rho}))  \\  
  \text{one-hop}  &= \sqrt{\frac{d_{\rho}^2}{9} + (r_{\rho}\sin(\rho((1/3)\cdot 0 + 0)) - r_{\rho}\sin(\rho((1/3)\cdot 0 + 1)+\frac{2\pi}{3}))^2  +
    (r_{\rho}\cos(\rho((1/3)\cdot0+0)) - r_{\rho}\cos(\rho((1/3)\cdot 0+1) + \frac{2\pi}{3})))^2} \\
  \text{one-hop}  &= \sqrt{\frac{d_{\rho}^2}{9} + (0  - r_{\rho}\sin(\rho/3 + \frac{2\pi}{3}))^2  +
    (r_{\rho} - r_{\rho}\cos(\rho/3 + \frac{2\pi}{3}))^2} \\
  \text{one-hop}  &= \sqrt{\frac{d_{\rho}^2}{9} + r_{\rho}^2\sin^2(\rho/3 + \frac{2\pi}{3})  +
    r_{\rho}^2(1 - \cos(\rho/3 + \frac{2\pi}{3}))^2} \\
  \text{one-hop}  &= \sqrt{\frac{d_{\rho}^2}{9} + r_{\rho}^2(\sin^2(\rho/3 + \frac{2\pi}{3})  + (1 - \cos(\rho/3 + \frac{2\pi}{3}))^2)} \\
  \text{but... } & d_{\rho}^2 &= 1 - 4 r_{\rho}^2 (\sin{ \rho / 2})^2 \\
  \text{one-hop}  &= \sqrt{\frac{1 - 4 r_{\rho}^2 (\sin{ \rho / 2})^2}{9} + r_{\rho}^2(\sin^2(\rho + \frac{2\pi}{3})  + (1 - \cos(\rho + \frac{2\pi}{3}))^2)} \\  
\end{align*}

@@

By similar algebra and trigonometry:
\begin{align*}
  \text{two-hops}  &= \sqrt{4\frac{1 - 4 r_{\rho}^2 (\sin{ \rho / 2})^2}{9} + r_{\rho}^2(\sin^2(2\rho/3 + \frac{4\pi}{3})  + (1 - \cos(2\rho/3 + \frac{4\pi}{3}))^2)} \\
\end{align*}

@@

Let $f_{\rho} = \sin^2(\rho/3 + \frac{2\pi}{3})  + (1 - \cos(\rho/3 + \frac{2\pi}{3})^2$.
Let $g_{\rho} = \sin^2(2\rho/3 + \frac{4\pi}{3})  + (1 - \cos(2\rho/3 + \frac{4\pi}{3})^2$.

The expression $f_{\rho}$ is maximized when $\rho = 2\pi/6 = 60\degree$, so that
is in some sense the ``wildest'' rail angle which maximizes the lengths. Since this expression is increasing up until this angle,
we conclude that:
\begin{align*}
  \text{two-hops} - \text{one-hop}  &= \sqrt{\frac{4d_{\rho}^2}{9} + r_{\rho}^2g_{\rho}} - \sqrt{\frac{d_{\rho}^2}{9} + r_{\rho}^2f_{\rho}} 
\end{align*}

To Be Done: I need to plot this expression, see if it has a maximum, and see if it always decreasing in this range, to see if we can prove optimality.

Since $d_{\rho}$ and $r_{\rho}$ are positive, this difference decreases as $f_{\rho}$ increases. Therefore we decrease the minimax length
of the whole system as we increase $\rho$ up to $\pi/3$ up unto the point that the shorter, one-hop distance is equal to the rail-length ($1$).
Therefore, to optimize the whole system so long as $\rho < \pi/3$, we equate one-hop to $1$ to find the optimum radius:

\begin{align*}
  1 &= \sqrt{\frac{d_{\rho}^2}{9} + r_{\rho}^2(\sin^2(\rho/3 + \frac{2\pi}{3})  + (1 - \cos(\rho/3 + \frac{2\pi}{3}))^2)} \\
  r_{\rho} &= \frac{2 \sqrt{2} }{\sqrt{9 \sqrt{3} sin(\rho/3) + 9 \cos(\rho/3) + 2 cos(\rho)+ 16 }}
\end{align*}
(where $9 \cos(\rho/3) + 2 cos(\rho)+ 16  \neq 0$), which does not occur when $\rho \leq \pi/3$).

Therefore, by computing $r_{\rho}$ from this equation, we can construct minimax optimal tetrahelix for an $\rho < \pi/3$.

@@

\begin{theorem}
The equitetrabeam with minimal maximal edge difference is produced by $H_{general}$ when $ r = \sqrt{\frac{8}{27}} $.
  \end{theorem}

@@

We observe that it by thinking of the straight lines of the Equitetrabeam as a degenerate helix of zero curvature,
it should be possible to define a smoothly varying continuum between the Boerdijk--Coxeter helix and the Equitetrabeam and every
curvature and torsion between the two.

@@

\[
H_{continuum}(n,c,\lambda) =
\left [
  \begin{tabular}{c}
   $ r_{\lambda} \cos(\phi_{\lambda} \kappa + c 2 \pi /3) $\\
   $ r_{\lambda}  \sin(\phi_{\lambda} \kappa + c 2 \pi /3) $\\
   $ d_{\lambda} \kappa $
  \end{tabular}
  \right ],
\text{where:}
  \begin{tabular}{c}
    $\kappa = n + c/3$ \\
    $ \phi_{\lambda} =  \lambda \rho_{bc}  + 0 $\\
    $\rho_{bc} = (3 \theta - 2 \pi)$ \\
   $ \theta = \arccos(-2/3) $
  \end{tabular}      
\]
A value of $\lambda = 0$ generates the equitetrabeam, and $\lambda = 1$ generates the Boerdijk--Coxeter helix, and every
value $\lambda \in [-1,1]$ generates an attractive structure which if physically realized is an inherently rigid structure
with member lengths of no greater disparity than $16\%$.
In the sense of structural engineering, it would be a relatively strong space frame. When $\lvert \lambda \rvert > 1$, tetrahelices
are still produced, but as $\lvert \lambda \rvert >> 2$ these tetrahelices probably cease to be interesting to mechanical engineers,
because their ability to withstand axial forces seem to diminish significantly.

@@

>>>>

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newtheorem{conjecture}{Conjecture}
\newtheorem{proofsketch}{Proof Sketech}

\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}

\usepackage{draftwatermark}

@@

\title{Untwisting the Tetrahelix (v0.5)}
\author{Robert L. Read \texttt{read.robert@gmail.com} \and
  Robert Gatliff \texttt{robert@toubat.org}
}

@@

\begin{align*}
  \text{one-hop} &= dist(H_{general}(n,c,\rho,d_{\rho}),H_{general}(n,c+1,\rho,d_{\rho}),r_{\rho})  \\
  \text{one-hop} &= dist(H_{general}(0,0,\rho,d_{\rho}),H_{general}(0,1,\rho,d_{\rho},r_{\rho}))  \\  
  \text{one-hop}  &= \sqrt{\frac{d_{\rho}^2}{9} + (r_{\rho}\sin(\rho((1/3)\cdot 0 + 0)) - r_{\rho}\sin(\rho((1/3)\cdot 0 + 1)+\frac{2\pi}{3}))^2  +
    (r_{\rho}\cos(\rho((1/3)\cdot0+0)) - r_{\rho}\cos(\rho((1/3)\cdot 0+1) + \frac{2\pi}{3})))^2} \\
  \text{one-hop}  &= \sqrt{\frac{d_{\rho}^2}{9} + (0  - r_{\rho}\sin(\rho/3 + \frac{2\pi}{3}))^2  +
    (r_{\rho} - r_{\rho}\cos(\rho/3 + \frac{2\pi}{3}))^2} \\
  \text{one-hop}  &= \sqrt{\frac{d_{\rho}^2}{9} + r_{\rho}^2\sin^2(\rho/3 + \frac{2\pi}{3})  +
    r_{\rho}^2(1 - \cos(\rho/3 + \frac{2\pi}{3}))^2} \\
  %% Formula above checks at sqrt(8/27), 1, 0, and rho_bc, h_bc, r_bc.
  \text{one-hop}  &= \sqrt{\frac{d_{\rho}^2}{9} + r_{\rho}^2(\sin^2(\rho/3 + \frac{2\pi}{3})  + (1 - \cos(\rho/3 + \frac{2\pi}{3}))^2)} \\
  d_{\rho}^2 &= 1 - 4 r_{\rho}^2 (\sin( \rho / 2)^2 \\
  %% Formula above checks at sqrt(8/27), 1, 0, and rho_bc, h_bc, r_bc.  
  \text{one-hop}  &= \sqrt{\frac{1}{9}  + r_{\rho}^2(-\frac{4 (\sin^2( \rho / 2))}{9} + \sin^2(\rho/3+ \frac{2\pi}{3})  + (1 - \cos(\rho/3 + \frac{2\pi}{3}))^2)} \\
  %% Formula above checks at sqrt(8/27), 1, 0, and rho_bc, h_bc, r_bc.    
\end{align*}

@@

By similar algebra and trigonometry:
\begin{align*}
  \text{two-hops}  &= \sqrt{\frac{4}{9} + r_{\rho}^2 (-\frac{16 (\sin^2( \rho / 2))}{9} + (\sin^2(2\rho/3 + \frac{4\pi}{3})  + (1 - \cos(2\rho/3 + \frac{4\pi}{3}))^2)} \\
\end{align*}

@@

Let $f_{\rho} = -\frac{4 (\sin^2( \rho / 2))}{9} + \sin^2(\rho/3+ \frac{2\pi}{3})  + (1 - \cos(\rho/3 + \frac{2\pi}{3}))^2)$.
Let $g_{\rho} = -\frac{16 (\sin^2( \rho / 2))}{9} + (\sin^2(2\rho/3 + \frac{4\pi}{3})  + (1 - \cos(2\rho/3 + \frac{4\pi}{3}))^2)$.

\begin{align*}
  \text{two-hops} - \text{one-hop}  &= \sqrt{\frac{4}{9} + r_{\rho}^2g_{\rho}} - \sqrt{\frac{1}{9} + r_{\rho}^2f_{\rho}} 
\end{align*}

If the partial derivative of $\text{two-hops} - \text{one-hop}$ with respect to the
radius is negative up until $\rho$ up until $\rho_{bc}$ when it is $0$, then
we will opitmize the overall minimax distance by choosing the largest radius
up until one-hop $= 1$, the rail-edge length. Numerical test suggests that this is true.

Therefore we decrease the minimax length
of the whole system as we increase the radius
up unto the point that the shorter, one-hop distance is equal to the rail-length ($1$).
Therefore, to optimize the whole system so long as $\rho < \rho_{bc}$,
we equate one-hop to $1$ to find the optimum radius:

\begin{align*}
  1 &=  \sqrt{\frac{1}{9}  + r_{\rho}^2(-\frac{4 (\sin^2( \rho / 2))}{9} + \sin^2(\rho/3+ \frac{2\pi}{3})  + (1 - \cos(\rho/3 + \frac{2\pi}{3}))^2)} \\
  r_{opt} &= \frac{2}{\sqrt{\frac{9}{2} \cdot ( \sqrt{3} sin(\rho/3) + \cos(\rho/3)) + cos(\rho)+ 8 }} \numberthis  \label{eqrhoopt} \\
\end{align*}
(where denominator $  \neq 0$).

Note: $ d_{opt} $ is computed by $\rho, r_{opt}$ via the rail equation \eqref{eqrhoopt}.

Therefore, by computing $r_{opt}$ as a function of $\rho$ from this equation, we can construct minimax optimal tetrahelix for an $\rho < \rho_{bc}$.

\begin{theorem}
  \label{optimality}
  For any rail angle $\rho$ where $\rho \leq \rho_{bc}$, a tetrahelix of minimum maximum edge length
  difference is generated by $H_{general}(n,c,\rho,d_{\rho},r_{opt})$, where $r_{opt}$ and $d_{opt}$ are
  computed from $\rho$. 
\end{theorem}

\begin{proof}
  TBD: The Even thirds theorem tells us an optimal will have even thirds.  We have shown
  that $r_{opt}$ minimizes the maximum length.
  \end{proof}

@@

\begin{corollary}
  The equitetrabeam with minimal maximal edge difference is produced
  by $H_{general}$ when $ r = \sqrt{\frac{8}{27}} $.
  \end{corollary}

@@

We observe that by thinking of the straight lines of the
Equitetrabeam as a degenerate helix of zero curvature,
it should be possible to define a smoothly varying continuum between the Boerdijk--Coxeter helix and the Equitetrabeam and every
curvature and torsion between the two.

@@

\[
H_{continuum}(n,c,\lambda) =
\left [
  \begin{tabular}{c}
   $ r_{\lambda} \cos(\phi_{\lambda} \kappa + c 2 \pi /3) $\\
   $ r_{\lambda}  \sin(\phi_{\lambda} \kappa + c 2 \pi /3) $\\
   $ d_{\lambda} \kappa $
  \end{tabular}
  \right ],
\text{where:}
  \begin{tabular}{c}
    $\kappa = n + c/3$ \\
    $ r_{opt} =  \text{is used : \eqref{eqrhoopt}} $ \\    
    $ \phi_{\lambda} =  \lambda \rho_{bc}  + 0 $\\
    $\rho_{bc} = (3 \theta - 2 \pi)$ \\
   $ \theta = \arccos(-2/3) $
  \end{tabular}      
\]
A value of $\lambda = 0$ generates the equitetrabeam, and $\lambda = 1$ generates the Boerdijk--Coxeter helix, and every
value $\lambda \in [-1,1]$ generates an attractive structure which if physically realized is an inherently rigid structure
with member lengths of no greater disparity than $16\%$.
In the sense of structural engineering, it would be a relatively strong space frame. When $\lvert \lambda \rvert > 1$, tetrahelices
are still produced, but as $\lvert \lambda \rvert >> 2$ these tetrahelices probably cease to be interesting to mechanical engineers,
because their ability to withstand axial forces seem to diminish significantly.

@@

Check every step of my derivation at rho = 0 and the BC values.

Found something.  So now the computation of the optimal radii seems slightly better than what we had before.
We have not, however, proven it to be optimal in the ranges involved, but at least we seem to be computing it correctly.
